---
title: "power-analysis"
author: "Zachary Houghton"
date: "2025-01-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(brms)
library(MASS)  # For mvrnorm()
```

# Power Analysis

In this markdown we'll demonstrate a power analysis for an experiment that is examining whether the less frequent form of a binomial primes the more frequent form of the binomial even more than it does the less frequent form of the binomial.

The details can be found in [link paper and pre-registration].

Here is the data scheme and task. Each participant took part of a self-paced reading task with a prime sentence and a target sentence.

For each row in our data we have:

-   residual reading time/rt (dv)

-   prime_freq: whether the prime was for the not-frequent ordering of the binomial or whether it was unrelated

-   target_freq: whether the target sentence that the participant was reading contained the frequent or not-frequent ordering of the binomial

The hypothesis was that due to noisy-channel processing *butter and bread* may speed up the processing of *bread and butter* even faster than it speeds up the processing of *butter and bread.* In other words, the speed up of the not-frequent prime may be larger for the frequent target than the not-frequent target.

Let's jump into the power analysis. We are generating data assuming the following hierarchical structure:

rt = (b0 + b0j + b0k) + (b1 + b1j + b1k) \* prime_freq + (b2 + b2j + b2k) \* target_freq + (b3 + b3j + b3k) \* (prime_freq \* target_freq) + epsiloni )

$$
rt \sim (\beta_0 + \beta_{0j} + \beta_{0k}) + (\beta_1 + \beta_{1j} + \beta_{1k}) \times prime\_freq + (\beta_2 + \beta_{2j} + \beta_{2k}) \times target\_freq + (\beta_1 + \beta_{1j} + \beta_{1k}) \times (prime\_freq * target\_freq) + \epsilon_i
$$

A pretty standard setup for psycholinguistics. We are also going to treatment code our variables such that the unrelated prime is 0 and the not-frequent prime is 1, and the not-frequent target is 0 and the frequent target is 1. This way we can focus on whether the interaction effect is less than zero.

## Data Generation

First let's load our data:

```{r}
data = read_csv('../Data/data_for_power.csv') %>% #pilot data
  dplyr::select('Item', 'prime_freq', 'target_freq')
head(data,3)
```

I ran a pilot version of this experiment and ran a Bayesian mixed-effects model to get each of these values. If I hadn't, I would have used values from a similar study in the literature. However, I prefer to pilot an experiment and get estimates that way, it feels more elegant.

First let's get the coefficient estimates and the subject-level sd and correlations for each term. Notice that there are 4 terms, so there are 6 different correlations (4 choose 2).

```{r}

b0 = -49.54 #Intercept
b1 = -7.46 #prime
b2 = -4.79 #target
b3 = -4.42 #interaction

#standard deviations from pilot data
#these are our subject-level effects 
sd_b0j = 115  #Intercept
sd_b1j = 17   #Prime Frequency
sd_b2j = 30  #Target Frequency
sd_b3j = 25  #Interaction (Prime*Target)

#correlations from pilot brms model
cor_b0j_b1j = 0.11    #Intercept,Prime Frequency
cor_b0j_b2j = -0.29   #Intercept,Target Frequency
cor_b0j_b3j = -0.15   #Intercept,prime:target
cor_b1j_b2j = -0.02    #Prime Frequency,Target Frequency
cor_b1j_b3j = -0.05   #Prime Frequency,Interaction
cor_b2j_b3j = 0.06    #Target Frequency,Interaction

```

The next part can be a bit tricky. We need to create a covariance-variance matrix because when we generate our subject-level random effects, we're going to sample them from a multivariate distribution with correlations between our random-effects and correlations between the intercepts and slopes. We can do this by remembering a bit about the structure of covariance-variance matrices.

There are two important points:

First, the covariance of two coefficients is equal to the correlation between the two, times the standard deviation of each. In more mathematical terms:

$$
cov(x,y) = cor(x,y) \times sd(x) \times sd(y)
$$

Second, recall that covariance-variance matrices follow the following format:

$$
\begin{bmatrix}
\text{Var}(x_1) & \cdots & \text{Cov}(x_1, x_n) \\
\vdots & \ddots & \vdots \\
\text{Cov}(x_n, x_1) & \cdots & \text{Var}(x_n)
\end{bmatrix}
$$

Once we recognize these two points, it becomes clear that we have all the information already to make the matrix. After this we can just sample from a multivariate distribution using it.

Our initial dataset will assume 25 participants.

```{r}
#variance-covariance matrix for subjects
#this part can be a bit tricky to follow
subject_varcov = matrix(c(
#here's the format:
  #Var(x1) ... Cov(x1,xn)
  #cov(xn,x1) ... var(xn)
  #we do this for both items and subjects
  
  sd_b0j^2,  cor_b0j_b1j * sd_b0j * sd_b1j,  cor_b0j_b2j * sd_b0j * sd_b2j,  cor_b0j_b3j * sd_b0j * sd_b3j,
  cor_b0j_b1j * sd_b0j * sd_b1j,  sd_b1j^2,  cor_b1j_b2j * sd_b1j * sd_b2j,  cor_b1j_b3j * sd_b1j * sd_b3j,
  cor_b0j_b2j * sd_b0j * sd_b2j,  cor_b1j_b2j * sd_b1j * sd_b2j,  sd_b2j^2,  cor_b2j_b3j * sd_b2j * sd_b3j,
  cor_b0j_b3j * sd_b0j * sd_b3j,  cor_b1j_b3j * sd_b1j * sd_b3j,  cor_b2j_b3j * sd_b2j * sd_b3j,  sd_b3j^2
), nrow = 4, byrow = TRUE)

#correlated subject effects
n_subjects = 25 #we'll change this number to see how many subjects we need
subject_effects = mvrnorm(n_subjects, mu = c(0, 0, 0, 0), Sigma = subject_varcov)

# Assign each column to corresponding effect
b0j = subject_effects[,1]  # Subject-specific intercepts
b1j = subject_effects[,2]  # Subject-specific prime_freq slopes
b2j = subject_effects[,3]  # Subject-specific target_freq slopes
b3j = subject_effects[,4]  # Subject-specific interaction slopes
noise_sd = 241
```

Next we can do the same for items:

```{r}
#Item level effects
#standard deviations from pilot data
sd_b0k = 25  #Intercept
sd_b1k = 17  #Prime Frequency
sd_b2k = 25  #Target Frequency
sd_b3k = 23  #Interaction (Prime*Target)

#correlations for items
cor_b0k_b1k = -0.16   # Intercept,Prime 
cor_b0k_b2k = -0.12   # Intercept,Target 
cor_b0k_b3k = -0.13   # Intercept,Interaction
cor_b1k_b2k =  -0.00   # Prime,Target
cor_b1k_b3k = 0.02   # Prime,Prime:target
cor_b2k_b3k = -0.01   # Target,prime:target

# Create variance-covariance matrix for items
item_varcov = matrix(c(
  sd_b0k^2,  cor_b0k_b1k * sd_b0k * sd_b1k,  cor_b0k_b2k * sd_b0k * sd_b2k,  cor_b0k_b3k * sd_b0k * sd_b3k,
  cor_b0k_b1k * sd_b0k * sd_b1k,  sd_b1k^2,  cor_b1k_b2k * sd_b1k * sd_b2k,  cor_b1k_b3k * sd_b1k * sd_b3k,
  cor_b0k_b2k * sd_b0k * sd_b2k,  cor_b1k_b2k * sd_b1k * sd_b2k,  sd_b2k^2,  cor_b2k_b3k * sd_b2k * sd_b3k,
  cor_b0k_b3k * sd_b0k * sd_b3k,  cor_b1k_b3k * sd_b1k * sd_b3k,  cor_b2k_b3k * sd_b2k * sd_b3k,  sd_b3k^2
), nrow = 4, byrow = TRUE)

# Simulate correlated item effects
n_items = length(data$Item)
item_effects = mvrnorm(n_items, mu = c(0, 0, 0, 0), Sigma = item_varcov)

# Assign each column to corresponding effect
b0k = item_effects[,1]  # Item-specific intercepts
b1k = item_effects[,2]  # Item-specific prime_freq slopes
b2k = item_effects[,3]  # Item-specific target_freq slopes
b3k = item_effects[,4]  # Item-specific interaction slopes
```

And finally we generate our data:

```{r}
n = n_subjects  # Number of subjects

d = tibble(
  subject = rep(1:n, each = n_items),
  item = rep(1:n_items, times = n)
) %>%
  mutate(
    # Assign subject and item effects
    b0j = rep(b0j, each = n_items),
    b1j = rep(b1j, each = n_items),
    b2j = rep(b2j, each = n_items),
    b3j = rep(b3j, each = n_items),
    b0k = rep(b0k, times = n),
    b1k = rep(b1k, times = n),
    b2k = rep(b2k, times = n),
    b3k = rep(b3k, times = n),
    
    # Error term
    epsiloni = rnorm(n * n_items, mean = 0, sd = noise_sd),
    
    # Condition coding
    prime_freq = rep(ifelse(data$prime_freq == 'unrelated', 0, 1), times = n),
    target_freq = rep(ifelse(data$target_freq == 'not-frequent', 0, 1), times = n),
    
    # Response time computation
    rt = (b0 + b0j + b0k) + 
         (b1 + b1j + b1k) * prime_freq + 
         (b2 + b2j + b2k) * target_freq + 
         (b3 + b3j + b3k) * (prime_freq * target_freq) + 
         epsiloni
  )


d$subject = factor(d$subject)
d$item = factor(d$item)
d$target_freq = factor(d$target_freq)
d$prime_freq = factor(d$prime_freq)
```

## Power Simulations

Now we use our previous data generation process and we put it into a function.

```{r}
sim_d = function(seed, n) {
  
  set.seed(seed)
  
  n_subjects = n

  
  b0 = -34 #Intercept
  b1 = -15 #prime
  b2 = -6 #target
  b3 = -12 #interaction

  #standard deviations from pilot data
  #these are our subject-level effects 
  b0j = 33  #Intercept
  b1j = 10   #Prime Frequency
  b2j = 10  #Target Frequency
  b3j = 10  #Interaction (Prime*Target)

  
  #standard deviations from pilot data
  #these are our subject-level effects 
  sd_b0j = 33  #Intercept
  sd_b1j = 10   #Prime Frequency
  sd_b2j = 10  #Target Frequency
  sd_b3j = 10  #Interaction (Prime*Target)
  
  #correlations from pilot brms model
  cor_b0j_b1j = 0.11    #Intercept,Prime Frequency
  cor_b0j_b2j = -0.29   #Intercept,Target Frequency
  cor_b0j_b3j = -0.15   #Intercept,prime:target
  cor_b1j_b2j = -0.02    #Prime Frequency,Target Frequency
  cor_b1j_b3j = -0.05   #Prime Frequency,Interaction
  cor_b2j_b3j = 0.06    #Target Frequency,Interaction
  
  #variance-covariance matrix for subjects
  #this part was hard lol
  subject_varcov = matrix(c(
  #here's the format:
    #Var(x1) ... Cov(x1,xn)
    #cov(xn,x1) ... var(xn)
    #we do this for both items and subjects
    
    sd_b0j^2,  cor_b0j_b1j * sd_b0j * sd_b1j,  cor_b0j_b2j * sd_b0j * sd_b2j,  cor_b0j_b3j * sd_b0j * sd_b3j,
    cor_b0j_b1j * sd_b0j * sd_b1j,  sd_b1j^2,  cor_b1j_b2j * sd_b1j * sd_b2j,  cor_b1j_b3j * sd_b1j * sd_b3j,
    cor_b0j_b2j * sd_b0j * sd_b2j,  cor_b1j_b2j * sd_b1j * sd_b2j,  sd_b2j^2,  cor_b2j_b3j * sd_b2j * sd_b3j,
    cor_b0j_b3j * sd_b0j * sd_b3j,  cor_b1j_b3j * sd_b1j * sd_b3j,  cor_b2j_b3j * sd_b2j * sd_b3j,  sd_b3j^2
  ), nrow = 4, byrow = TRUE)
  all(subject_varcov == t(subject_varcov)) # check that it's a valid variance-covariance matrix
  #correlated subject effects

  subject_effects = mvrnorm(n_subjects, mu = c(0, 0, 0, 0), Sigma = subject_varcov)
  
  # Assign each column to corresponding effect
  b0j = subject_effects[,1]  # Subject-specific intercepts
  b1j = subject_effects[,2]  # Subject-specific prime_freq slopes
  b2j = subject_effects[,3]  # Subject-specific target_freq slopes
  b3j = subject_effects[,4]  # Subject-specific interaction slopes
  
  noise_sd = 100
  
  #Item level effects
  #standard deviations from pilot data
  sd_b0k = 9  #Intercept
  sd_b1k = 10  #Prime Frequency
  sd_b2k = 10  #Target Frequency
  sd_b3k = 10  #Interaction (Prime*Target)
  
  #correlations for items
  cor_b0k_b1k = -0.16   # Intercept,Prime 
  cor_b0k_b2k = -0.12   # Intercept,Target 
  cor_b0k_b3k = -0.13   # Intercept,Interaction
  cor_b1k_b2k =  -0.00   # Prime,Target
  cor_b1k_b3k = 0.02   # Prime,Prime:target
  cor_b2k_b3k = -0.01   # Target,prime:target
  
  # Create variance-covariance matrix for items
  item_varcov = matrix(c(
    sd_b0k^2,  cor_b0k_b1k * sd_b0k * sd_b1k,  cor_b0k_b2k * sd_b0k * sd_b2k,  cor_b0k_b3k * sd_b0k * sd_b3k,
    cor_b0k_b1k * sd_b0k * sd_b1k,  sd_b1k^2,  cor_b1k_b2k * sd_b1k * sd_b2k,  cor_b1k_b3k * sd_b1k * sd_b3k,
    cor_b0k_b2k * sd_b0k * sd_b2k,  cor_b1k_b2k * sd_b1k * sd_b2k,  sd_b2k^2,  cor_b2k_b3k * sd_b2k * sd_b3k,
    cor_b0k_b3k * sd_b0k * sd_b3k,  cor_b1k_b3k * sd_b1k * sd_b3k,  cor_b2k_b3k * sd_b2k * sd_b3k,  sd_b3k^2
  ), nrow = 4, byrow = TRUE)
  #all(item_varcov == t(item_varcov)) # check that it's a valid variance-covariance matrix
  # Simulate correlated item effects
  n_items = length(data$Item)
  item_effects = mvrnorm(n_items, mu = c(0, 0, 0, 0), Sigma = item_varcov)
  
  # Assign each column to corresponding effect
  b0k = item_effects[,1]  # Item-specific intercepts
  b1k = item_effects[,2]  # Item-specific prime_freq slopes
  b2k = item_effects[,3]  # Item-specific target_freq slopes
  b3k = item_effects[,4]  # Item-specific interaction slopes
  
  d = tibble(
    subject = rep(1:n, each = n_items),
    item = rep(1:n_items, times = n)
  ) %>%
    mutate(
      # Assign subject and item effects
      b0j = rep(b0j, each = n_items),
      b1j = rep(b1j, each = n_items),
      b2j = rep(b2j, each = n_items),
      b3j = rep(b3j, each = n_items),
      b0k = rep(b0k, times = n),
      b1k = rep(b1k, times = n),
      b2k = rep(b2k, times = n),
      b3k = rep(b3k, times = n),
      
      # Error term
      epsiloni = rnorm(n * n_items, mean = 0, sd = noise_sd),
      
      # Condition coding
      prime_freq = rep(ifelse(data$prime_freq == 'unrelated', 0, 1), times = n),
      target_freq = rep(ifelse(data$target_freq == 'not-frequent', 0, 1), times = n),
      
      # Response time computation
      rt = (b0 + b0j + b0k) + 
           (b1 + b1j + b1k) * prime_freq + 
           (b2 + b2j + b2k) * target_freq + 
           (b3 + b3j + b3k) * (prime_freq * target_freq) + 
           epsiloni
    )
  
  
  d$subject = factor(d$subject)
  d$item = factor(d$item)
  d$target_freq = factor(d$target_freq)
  d$prime_freq = factor(d$prime_freq)
  
 d = d #so that d is returned
}

```

Next we fit a single model. We do this so that we can use brms update() function to hopefully reduce some of the compilation time.

```{r}

## first model
t1 = Sys.time()
fit = 
  brm(data = d, cores = 4, chains = 4, iter = 6000, warmup = 3000,
      rt ~ 0 + Intercept + prime_freq*target_freq + 
        (prime_freq*target_freq|subject) + 
        (prime_freq*target_freq|item),
      #control = list(adapt_delta = 0.95),
      prior = c(set_prior("student_t(3, 0, 10)", class = "b"),
                set_prior("student_t(3, 0, 24.32)", class = "sd")), #from our model estimate
      seed = 465) #random seed
t2 = Sys.time()
t2 - t1 #11mins / 15.6mins'

```

Now we'll use a bit of tidyverse magic to run the same model across different generated data without having to compile the model over and over again. Courtesy of: <https://solomonkurz.netlify.app/blog/bayesian-power-analysis-part-i/>

We'll re-run the same model with 50 participants.

```{r}

n_sim = 100 #set to 2 initially for debugging

t1 = Sys.time()

s = #n = 100: power of 0.33
  tibble(seed = 1:n_sim) %>% 
  mutate(d = map(seed, sim_d, n = 500)) %>% #what about 100 participants?
  mutate(fit = map2(d, seed, ~update(fit, newdata = .x, seed = .y, cores = 4, chains = 4)))

t2 <- Sys.time()

t2 - t1
```

Notice how we don't have to re-compile the model. This saves us a ton of time. Though even just to run those two simulations, it took a `r t2 - t1`. So running 1000 simulations would take upwards of 33 hours. Not great, but I'm fortunate enough to be able to use university resources to run 1000 simulations. I'd probably only run 100 simulations if I didn't have a cpu cluster.

Finally we try different numbers of subjects to determine how many subjects we need.

```{r}
# compute power
n_sim = 100
n_subjs = 250

s = #n = 50, b0 = -20, b1 = 0: power of 0.33
  tibble(seed = 1:n_sim) %>% 
  mutate(d = map(seed, sim_d, n = n_subjs)) %>% 
  mutate(fit = map2(d, seed, ~update(fit, newdata = .x, seed = .y, 
                                     chains = 4, cores = 4, iter = 4000, warmup = 2000)))

```

Now we can compute the proportion of models where the coefficient estimate credible intervals for our three effects of interest were less than zero. If I expected the effect to be positive we would look at whether the 2.5% interval was greater than zero instead.

```{r}
#summary statistics for fixed effects from each model
s = s %>%
  mutate(
    summary = map(fit, ~as.data.frame(fixef(.x, probs = c(0.025, 0.975)))), #Extract fixed effects
    prime_freq_check = map_lgl(summary, ~.x["prime_freq1", "Q97.5"] < 0), #Check if upper CI bound is < 0 for "prime_freq"
    target_freq_check = map_lgl(summary, ~.x["target_freq1", "Q97.5"] < 0),
    prime_target_check = map_lgl(summary, ~.x["prime_freq1:target_freq1", "Q97.5"] < 0),
    prime_freq = map_dbl(summary, ~.x["prime_freq1", "Estimate"]),
    target_freq = map_dbl(summary, ~.x["target_freq1", "Estimate"]),
    prime_target_freq = map_dbl(summary, ~.x["prime_freq1:target_freq1", "Estimate"]),
    prime_freq_2.5 = map_dbl(summary, ~.x["prime_freq1", "Q2.5"]),
    target_2.5 = map_dbl(summary, ~.x["target_freq1", "Q2.5"]),
    prime_target_2.5 = map_dbl(summary, ~.x["prime_freq1:target_freq1", "Q2.5"]),
    prime_freq_97.5 = map_dbl(summary, ~.x["prime_freq1", "Q97.5"]),
    target_97.5 = map_dbl(summary, ~.x["target_freq1", "Q97.5"]),
    prime_target_97.5 = map_dbl(summary, ~.x["prime_freq1:target_freq1", "Q97.5"])
  )

#proportion of models where "prime_freq" CI upper bound is < 0
prime_freq_below_zero = mean(s$prime_freq_check)
target_freq_below_zero = mean(s$target_freq_check)
prime_target_below_zero = mean(s$prime_target_check)

prime_2.5 = mean(s$prime_freq_2.5)
target_2.5 = mean(s$target_2.5)
prime_target_2.5 = mean(s$prime_target_2.5)

prime_97.5 = mean(s$prime_freq_97.5)
target_97.5 = mean(s$target_97.5)
prime_target_97.5 = mean(s$prime_target_97.5)

prime_est = mean(s$prime_freq)
target_est = mean(s$target_freq)
prime_target_est = mean(s$prime_target_freq)

prime_freq_below_zero
target_freq_below_zero
prime_target_below_zero

prime_2.5 
target_2.5
prime_target_2.5

prime_97.5
target_97.5
prime_target_97.5
```

```{r}

```
